---
title: "ICT583 Assignment 2"
author: "Supuni Agampodi"
date: "2023-05-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The dataset used in this project was collected from mobile health care services provided in collaboration with elderly care centers operated by local non-governmental organizations. These health care services were offered to elderly individuals residing in different districts of Hong Kong. The services were provided free of charge over a ten-year period, specifically from 2008 to 2018. The dataset likely contains information related to the health status, demographics, and other relevant factors of the community-dwelling elderly population in Hong Kong during that time frame.

```{r}
setwd("D:/Supuni/Murdoch/Y2S2/ICT515-DS/Assignments/Assignment 2")
data <- read.csv("dataset ICT583 2023.csv", header = TRUE, sep = ",")
dim(data)
head(data)
tail(data)

str(data)
table(data$MMSE_class_binary)
pairs(data[,2:13], lower.panel = NULL)

```


## Understand Data - Summary Statistics

The summary statistics before data pre processing.


```{r}

names(data)

# Continuous variables
continuous <- c("Age", "Body_Height", "Body_Weight", "MNAa_total", "MNAb_total", "GDS")

# Categorical variables
categorical <- c("Gender", "Education_ID", "Financial_status", "Independent_or_depend_on_family",
                 "Marital_status_ID", "MMSE_class_binary")

data %>% select(all_of( categorical)) %>%  map(table)

data %>% select(all_of( continuous)) %>%
  map(., ~{
    
    c(
    Mean = mean(.x, na.rm = T) %>% round(2),
    SD = sd(.x, na.rm = T)%>% round(2)
    )
    
  }
)

# missing values
data %>% is.na %>% colSums()
sum(is.na(data))

summary(data)

```


## Missing Values 

The total number of missing values for each variable was computed. Mean imputation and mode imputation [3] were done for continuous variables and categorical variables respectively.

```{r}

data %>% is.na %>% colSums()
sum(is.na(data))

# Impute missing values with mean value
data$MNAb_total[is.na(data$MNAb_total)] <- mean(data$MNAb_total, na.rm = TRUE)
data$MNAa_total[is.na(data$MNAa_total)] <- mean(data$MNAa_total, na.rm = TRUE)
data$GDS[is.na(data$GDS)] <- mean(data$GDS, na.rm = TRUE)

# # Impute missing values with mode 
# data$Financial_status[is.na(data$Financial_status)] <- mfv(data$Financial_status, method = "discrete")
# 
# data$Marital_status_ID[is.na(data$Marital_status_ID)] <- mfv(data$Marital_status_ID, method = "discrete")
# 
# data$Education_ID[is.na(data$Education_ID)] <- mfv(data$Education_ID, method = "discrete")
# 
# data$Independent_or_depend_on_family[is.na(data$Independent_or_depend_on_family)] <- mfv(data$Independent_or_depend_on_family, method = "discrete")

# Create a function to calculate the mode
calculate_mode <- function(x) {
  unique_x <- unique(x)
  tab_x <- table(x)
  mode_val <- unique_x[which.max(tab_x)]
  return(mode_val)
}

# Impute missing values with mode
data$Financial_status[is.na(data$Financial_status)] <- calculate_mode(data$Financial_status)
data$Marital_status_ID[is.na(data$Marital_status_ID)] <- calculate_mode(data$Marital_status_ID)
data$Education_ID[is.na(data$Education_ID)] <- calculate_mode(data$Education_ID)
data$Independent_or_depend_on_family[is.na(data$Independent_or_depend_on_family)] <- calculate_mode(data$Independent_or_depend_on_family)



data %>% select(all_of( categorical)) %>%  map(table)
view(data)
sum(is.na(data))
summary(data)
```
# Summary Statistic after missing value imputation
```{r}

data %>% select(all_of( categorical)) %>%  map(table)

data %>% select(all_of( continuous)) %>%
  map(., ~{
    
    c(
    Mean = mean(.x, na.rm = T) %>% round(2),
    SD = sd(.x, na.rm = T)%>% round(2)
    )
    
  }
)

```


# change class
```{r}

data1 <- data %>%
  mutate(across(all_of(categorical), as.factor))

```



# Corealtion
```{r}

corr = data %>% select_if(is.numeric) %>% 
  cor(., method = "spearman") %>% round(2) 

corr %>% write.csv("corr.csv") 

corr

```


# Dummy Variable 
```{r}

f1 = function(x) gsub(1, "yes", x) %>%
  gsub(0, "no", .) %>% as.character()

data = data %>% mutate(MMSE = f1(data$MMSE_class_binary))

str(data)

```

## change class

The total number of missing values for each variable was computed. No record was excluded because there are no missing values for MMSE_class_binary variable. Mean imputation and mode imputation [3] were done for continuous variables and categorical variables respectively.

```{r}

# Categorical variables
categorical1 <- c("Gender", "Education_ID", "Financial_status", "Independent_or_depend_on_family", "Marital_status_ID", "MMSE")

data <- data %>%mutate(across(all_of(categorical1), as.factor))

```


## Visualization

Boxplot, density plot and bar plot [4] were used to show difference in selected features between groups of MMSE.

```{r}
#plot(data, main = "Searching Outliers")
data1<-data
names(data)
ggsave("MMSE_Age.png", width = 5, height = 3)
ggplot(data=data1, aes(x=MMSE, y=Age))+geom_boxplot()

ggsave("MMSE_BH.png", width = 5, height = 3)
ggplot(data=data1, aes(x=MMSE, y=Body_Height))+geom_boxplot()

ggsave("MMSE_BW.png", width = 5, height = 3)
ggplot(data=data1, aes(x=MMSE, y=Body_Weight))+geom_boxplot()

ggsave("MMSE_GDS.png", width = 5, height = 3)
ggplot(data=data1, aes(x=MMSE, y=GDS))+geom_boxplot()

ggsave("MMSE_MNAa.png", width = 5, height = 3)
ggplot(data=data1, aes(x=MMSE, y=MNAa_total))+geom_boxplot()

ggsave("MMSE_MNAb.png", width = 5, height = 3)
ggplot(data=data1, aes(x=MMSE, y=MNAb_total))+geom_boxplot()

ggsave("MMSE_BWd.png", width = 5, height = 3)
ggplot(data, aes(Body_Weight, fill = MMSE)) + 
  geom_density(alpha = .1)

ggsave("MMSE_BHd.png", width = 5, height = 3)
ggplot(data, aes(Body_Height, fill = MMSE)) + 
  geom_density(alpha = .1)

ggsave("MMSE_MNAbd.png", width = 5, height = 3)
ggplot(data, aes(MNAa_total, fill = MMSE)) + 
  geom_density(alpha = .1)

ggsave("MMSE_MNAbd.png", width = 5, height = 3)
ggplot(data, aes(MNAb_total, fill = MMSE)) + 
  geom_density(alpha = .1)

```
```{r}
names(data)

ggsave("MMSE_Education.png", width = 5, height = 3)
ggplot(data, aes(MMSE, fill = Education_ID)) + 
  geom_bar(position = "dodge")


ggsave("MMSE_Gender.png", width = 5, height = 3)
ggplot(data, aes(MMSE, fill = Gender)) + 
  geom_bar(position = "dodge")


ggsave("MMSE_Financial_Status.png", width = 5, height = 3)
ggplot(data, aes(MMSE, fill = Financial_status)) + 
  geom_bar(position = "dodge")


ggsave("MMSE_Marital_Status.png", width = 5, height = 3)
ggplot(data, aes(MMSE, fill = Marital_status_ID)) + 
  geom_bar(position = "dodge")


ggsave("MMSE_Dependency.png", width = 5, height = 3)
ggplot(data, aes(MMSE, fill = Independent_or_depend_on_family)) + 
  geom_bar(position = "dodge")
```


## Scaling
Centering and scaling were done for continuous variables, after which they have mean = 0 and SD = 1.


```{r}

data2 <- data %>% mutate_if(is.numeric, scale)
#data2 = data %>% mutate_if(~ n_distinct(.) > 2, scale)

str(data2)

```

## Partitioning
The preprocessed data were randomly split into training dataset (70%) and testing dataset (30%), with a seed number set.

```{r}

set.seed(1)
dim(data2)
id <- createDataPartition(data2$MMSE, p = .7, list = F)

training <- data2[ id,]

testing <- data2[-id,]

dim(training)
dim(testing)

```

#Formula
The conceptual framework of the project is presented as a formula for modeling.


```{r}
names(data2)
form1 = MMSE ~ Age +  Gender + + Education_ID + GDS + Body_Height + MNAb_total + Independent_or_depend_on_family

form1
```
##Logistic regression
A logistic regression model was built, and the odds ratios were computed for the estimated log odds.

```{r}
str(data2)
sum(is.na(training))

logit_model<-glm(form1 , data = data2, family = binomial(logit))


s = summary(logit_model)
s

lr_coef = s $ coefficients %>% as.data.frame() %>%
  mutate( odds_ratio = exp(Estimate)) %>% round(3)

lr_coef %>% write.csv("lr_coef.csv")

lr_coef
```

##Models
The logistic regression classifier and SVM classifier [5] were trained and used to generate predicted classes and probabilities. The performance metrics were computed based on the confusion matrix [6]. ROC curves of the models were compared [7].

```{r}

logit_prob = predict(logit_model, testing, "response")

str(logit_prob)

logit_pred <- round(predict(LGM, data2, type="response"))


str(logit_pred)

```

```{r}
# Assuming you have a logistic regression model named logit_model

# Predict probabilities
logit_prob <- predict(logit_model, newdata = testing, type = "response")

# Check the structure of logit_prob
str(logit_prob)

# Round the predicted probabilities to get class predictions (0 or 1)
logit_pred <- ifelse(logit_prob > 0.5, 1, 0)

# Check the structure of logit_pred
str(logit_pred)

```



```{r}

#logit_perf = confusionMatrix(table(logit_pred, data[,"MMSE_class_binary"]))
#logit_perf

# Calculate the confusion matrix and performance metrics
logit_perf <- confusionMatrix(data = logit_pred, reference = data[,"MMSE_class_binary"])

# Print the performance metrics
logit_perf
```



# ROC - Logistic Regression
```{r}
lr_roc = roc(testing$MMSE, logit_prob)
lr_roc

```


# Support Vector Machine
```{r}
svm_model = svm(form1, training, probability = T)

summary(svm_model)

svm_pred = predict(svm_model, testing)

str(svm_pred)

svm_prob = predict(svm_model, testing, probability = T) %>%
  attr("probabilities") %>% .[,1]


str(svm_prob)


```




# SVM - Confusion Matrix
```{r}

#svm_perf = table(svm_pred, testing $ MMSE) %>% 
 # confusionMatrix(., positive = "yes")

svm.table = table(svm_pred,testing$MMSE)

svm_perf = confusionMatrix(svm.table)
svm_perf

```


# ROC -SVM
```{r}
svm_roc = roc(testing$MMSE, svm_prob)
svm_roc

```

# Performance Metrix
```{r}

perf_metrices=
cbind.data.frame(
  lr = c(logit_perf$overall, logit_perf$byClass, AUC = lr_roc$auc),
  svm = c(svm_perf$overall, svm_perf$byClass, AUC = svm_roc$auc)
  ) %>%
  round(3) %>%
  mutate(best = apply(., 1, which.max)) 

perf_metrices%>%
  write.csv("perf_metrices.csv")

perf_metrices

```

# ROC compare
```{r}

ggsave("roc.png", width = 6.5, height = 6.5)


ggroc(
  list(lr_roc, svm_roc),
  legacy.axes = TRUE ) +
  scale_color_discrete(
    labels = c(
      paste("Logistic Regression, AUC", round(lr_roc$auc, 3)),
      paste("Support vector Machine, AUC", round(svm_roc$auc, 3))
    )
  ) +
  labs(col = "Model") +
  theme_classic() +
  theme(legend.position = c(.8,.2))
```

## Reference List

Reference List
[1]   	R Core Team, "R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria," http://www. R-project. org/, 2016.
[2]   	RStudio Team, "RStudio: integrated development for R. RStudio," Inc., Boston, MA, vol. 700, p. 879, 2015.
[3]   	missMethods: Methods for Missing Data. R package version 0.4.0. (2022). [Online]. Available: https://CRAN.R-project.org/package=missMethods
[4]   	H. Wickham et al., "Welcome to the Tidyverse," Journal of open source software, vol. 4, no. 43, p. 1686, 2019.
[5]   	E. Dimitriadou, K. Hornik, F. Leisch, D. Meyer, and A. Weingessel, "Misc functions of the Department of Statistics (e1071), TU Wien," R package, vol. 1, pp. 5-24, 2008.
[6]   	M. Kuhn, "Building predictive models in R using the caret package," Journal of statistical software, vol. 28, pp. 1-26, 2008.
[7]   	X. Robin et al., "pROC: an open-source package for R and S+ to analyze and compare ROC curves," BMC bioinformatics, vol. 12, no. 1, pp. 1-8, 2011.
[8]	Wikipedia, "Geriatric Depression Scale," [Online]. Available: https://en.wikipedia.org/wiki/Geriatric_Depression_Scale, Accessed on: June 2, 2023.

